$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2014-03-31 JournalToSubmit2013/modelingWithPISODM.tex
Work on section 4: transformations

Given  (composite) $\pi$-use case model defining the functions of  the developed application and their associated constraints (representing business rules), the objective is to express them as a $\pi$-service process model describing the application logic and its associated non-functional properties.  The principle is to express a set of $\pi$-use cases (i.e., use cases with constraints) related by relationships of type {\sc Extend}   in terms of  a business process consisting of actions related by a control flow and associated contracts specifying NFP. The PIM to PIM transformation rules specify how  to transform a $\pi$-use case model into  a $\pi$-service process model.
%(Table \ref{fig:transformationsUseCase-ServiceProcess}).

@@

   A {\sf Constraint} of a (composite) {\sf Use Case}  is transformed into an  {\sf Assertion}.  The set of constraints  in a $\pi$-use case model  (transformed into assertions) are grouped in a {\sf Contract} in a $\pi$-service process model (figure  \ref{fig:transformationsUseCase-ServiceProcessRules}-a).
A constraint of a given type ({\sf Constraint type}) can have different transformations as stated in the following rules (figure   \ref{fig:transformationsUseCase-ServiceProcessRules}-b). Given a  a {\sf Constraint} of type:

\begin{itemize}
\item   {\sc Business} is transformed into an {\sf Assertion} with no value attributes. %information like maxValue and minValue;
\item     {\sc Value}  with the  attribute {\sf  isExceptionalBehaviour}  set to false, is transformed into an {\sf Assertion}.
\item  {\sc Value }  with the attribute {\sf  isExceptionalBehaviour }  set to true, is transformed into  {\sf Exceptional behaviour}.
\end{itemize}
%Given the $\pi$-UseCase meta-model, to all {\sc Constraint} entity related with a {\sc Use Case}, there is a {\sc Contract} that compounds a set of {\sc Assertions} entity  and the {\sc Use Case} is refined in a service {\sc Action}.

 For transforming constraints of type {\sf Value Constraint }
  the designer must specify value thresholds associated to attributes as specified by  { constraint}s  and {use case}s. By default, value constraints are transformed into pre-conditions and business constraints are transformed into post-conditions.

@@

The relationships of type  {\sc Extend} and {\sc Include}  determine the way the business process is expressed as a workflow.  The workflow generated for a relationship  «extends»  between  two use cases (figure  \ref{fig:transformationsUseCase-ServiceProcessRules}-a) is described in figure \ref{fig:transformaton-examples}-c).  The generated workflow consists of  a {\sf Fork node}, a {\sf Join node},  four entities of type {\sc Control flow}, and also two entities of type {\sc Action} , one for each {\sf Use case}.

When there is more than one «extends» relationship among several use cases (figures \label{fig:transformationsUseCase-ServiceProcessRuless}-b), the transformation leads to  two more {\sc Control flow} entities for each  {\sc use case}, and an {\sf Action} for each  extended {\sf use case}.

@@

 {\sf Include} use case entities, are transformed into an {\sf Action} sequence, as shown in figure \ref{fig:transformationsUseCase-ServiceProcessRules}-b. A {\sc Use case} element is transformed into an {\sf Action}. A set of n {\sc Use case} is transformed into an  n-1 {\sf Object flow} elements. Each {\sf Control flow} is transformed into two {\sf Actions} associated by a relationship.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT1}
Considering the scenario example  the transformation of the "listen music" Use Case is transformed into a Service Action. This Action  represents a Spotify service function that can be invoked to play the music. For the "publish music" use case,  constraints are transformed in a set of assertions that are grouped in a Contract ({\sf "publishMusicContract"}) associated to the Action {\sf "publishMusic"}. 
The "download music" use case  includes the payment process to buy the music. Thus, these use cases  are transformed into {\sf Actions}, and a {\sf Service activity} that aggregates these {\sf Actions}.   It is transformed in a sequence flow in the $\pi$-service process model (as shown in figure \ref{fig:transformation-example-include}). 
The same rule is applied for the "publish music" use case, which has two extended use cases, "public twitter" and "public Facebook" (figures \ref{fig:transformaton-examples}-b and d).
 \end{example}

@@

%Table \ref{fig:ServiceComposition-Pews-Rules} 
This section describes the PIM to PSM transformations between a $\pi$-Service composition and a $\pi$-PEWS program. We propose two groups of rules: those that transform services composition entities of a $\pi$-Service composition model into $\pi$-PEWS model ; and those that transform rules grouped by policies into A-policy types.

>>>>

Given  (composite) $\pi$-use case model defining the functions of  the developed application and  associated constraints (representing business rules), the objective is to express them as a $\pi$-service process model describing the application logic and its associated non-functional properties.  The principle is to express a set of $\pi$-use cases (i.e., use cases with constraints)   in terms of  a business process consisting of actions related by a control flow and associated contracts specifying NFP. The PIM to PIM transformation rules specify how  to transform a $\pi$-use case model into  a $\pi$-service process model.
%(Table \ref{fig:transformationsUseCase-ServiceProcess}).

@@

   A {\sf Constraint} of a (composite) {\sf Use Case}  is transformed into an  {\sf Assertion}.  The set of resulting assertions  are grouped in a {\sf Contract} in a $\pi$-service process model (figure  \ref{fig:transformationsUseCase-ServiceProcessRules}-a).
A constraint of a given type ({\sf Constraint type}) can have different transformations as stated in the following rules (figure   \ref{fig:transformationsUseCase-ServiceProcessRules}-b). Given a  a {\sf Constraint} of type:

\begin{itemize}
\item   {\sc Business} it is transformed into an {\sf Assertion} with no value attributes. %information like maxValue and minValue;
\item     {\sc Value}  with the  attribute {\sf  isExceptionalBehaviour}  set to false, is transformed into an {\sf Assertion}.
\item  {\sc Value }  with the attribute {\sf  isExceptionalBehaviour }  set to true, is transformed into  {\sf Exceptional behaviour}.
\end{itemize}
%Given the $\pi$-UseCase meta-model, to all {\sc Constraint} entity related with a {\sc Use Case}, there is a {\sc Contract} that compounds a set of {\sc Assertions} entity  and the {\sc Use Case} is refined in a service {\sc Action}.

 For transforming constraints of type {\sf Value Constraint }
  the designer must specify value thresholds associated to attributes as specified by  { constraint}s  in $\pi$-{use case}s. By default, value constraints are transformed into pre-conditions and business constraints are transformed into post-conditions.

@@

The relationships of type  {\sc Extend} and {\sc Include}  determine the way the business process is expressed as a workflow.  The workflow generated for a relationship  «extends»  between  two use cases (figure  \ref{fig:transformationsUseCase-ServiceProcessRules}-a) is described in figure \ref{fig:transformaton-examples}-c).  The generated workflow consists of  a {\sf Fork node}, a {\sf Join node},  four entities  {\sc Control flow}, and also two entities of type {\sc Action}  (one for each {\sf Use case}).

%When there is more than one «extends» relationship among  use cases (figures \label{fig:transformationsUseCase-ServiceProcessRuless}-b), the transformation leads to  two more {\sc Control flow} entities for each  {\sc use case}, and an {\sf Action} for each  extended {\sf use case}.

@@

 {\sf Include} use case entities, are transformed into an {\sf Action} sequence, as shown in figure \ref{fig:transformationsUseCase-ServiceProcessRules}-b. A {\sc Use case} element is transformed into an {\sf Action}. A set of n {\sc Use case} is transformed into an  n-1 {\sf Object flow} elements. 
 %Each {\sf Control flow} is transformed into two {\sf Actions} associated by a relationship.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT1}
Considering the scenario example  the transformation of the "listen music" Use Case is transformed into a Service Action. This Action  represents a Spotify service function that can be invoked to play the music. For the "publish music" use case,  constraints are transformed in a set of assertions that are grouped in a Contract ({\sf "publishMusicContract"}) associated to the Action {\sf "publishMusic"}. 
The "download music" use case  includes the payment process to buy the music. Thus, these use cases  are transformed into {\sf Actions}, and a {\sf Service activity} that aggregates these {\sf Actions}.   It is transformed in a sequence flow in the $\pi$-service process model.
%(as shown in figure \ref{fig:transformation-example-include}). 
The same rule is applied for the "publish music" use case, which has two extended use cases, "public twitter" and "public Facebook" (figures \ref{fig:transformaton-examples}-b and d).
 \end{example}

@@

%Table \ref{fig:ServiceComposition-Pews-Rules} 
This section describes the PIM to PSM transformations from a $\pi$-Service composition model to a $\pi$-PEWS program. We propose two groups of rules: those that transform services composition entities of a $\pi$-Service composition model into $\pi$-PEWS model; and those that transform rules grouped by policies into A-policy types.
